---
title: "Can sensory analysis predict wine better than chemical analysis?"
author: "Feng Oskar, Christian Schoenherr, Jose Sanchez Gavito"
date: "June 15, 2016"
output: html_document
---
<br>

The purpose of this project is to analyse two different sets that describe a group of wines;

1.-Based on the sensory description of the wines 
2.-Based on the chemical analysis of the wines


The established thesis is that winemaking is an art, and that sensory perception is a better instrument to identify the components that differentiate the taste of wines than the latest chemical instrumental analysis. The data set was used obtained from a study done at UC Davis, both the sensory and chemical analysis were done in triplicate to ensure reliable data.

The Analysis

A correlation matrix was done to determine the use of Factor Analysis and to better understand the main drivers of the data. The PCA were performed for both data sets to determine the principal components of the two first dimensions. 



```{r echo=TRUE, eval=FALSE, tidy=TRUE}
getwd()
setwd("CourseSessions/Sessions23/project")
list.files()
rm(list=ls()) # Clean up the memory, if we want to rerun from scratch
```



```{r echo=TRUE, eval=TRUE, tidy=TRUE}
sensorydata <- read.csv("data/sensorystand.csv", sep=",", dec=".")
sensory <- data.frame(sensorydata[,-1], row.names=sensorydata[,1])
round(cor(sensory[,1:15]), 2)
pcsensory<- princomp(sensory[,1:15], cor=TRUE, scores=TRUE)
summary(pcsensory)
plot(pcsensory,type="lines")
biplot(pcsensory)
```


The plot shows that the first dimension is positively dominated by Fruity and Tropical Aromas, and negatively correlated with spice, oak, hot MF (mouth feel) and bitterness

The second dimension is positively correlated to vegetable, hot, chem and sour while negatively correlated to sweet, caramel and viscous MF 

(All makes sense)

The overall space is well described by the attributes and some groupings of the wines can be seen in the picture. Chardonnays with similar taste (heavy oak influenced) Riesling with similar sweet character, Sauvignon Blanc with herbal and citrus and Viognier with fruity and tropical



```{r echo=TRUE, eval=TRUE, tidy=TRUE}
chemicaldata <- read.csv("data/chemicalstand.csv", sep=",", dec=".")
chemical <- data.frame(chemicaldata[,-1], row.names=sensorydata[,1])
round(cor(chemical[,1:16]), 2)
pcchemical<- princomp(chemical[,1:16], cor=TRUE, scores=TRUE)
summary(pcchemical)
plot(pcchemical,type="lines")
biplot(pcchemical)
```
<br>


The chemical PCA is a little more difficult to interpret. The vectors are not evenly distributed and a particular group of wines are not described by the chemistry (SauvB). The Chardonnay seems to be correlated with Glycerol (vanilla taste) makes sense for a Napa Chardonnay. Rieslings on the other hand are again described as sweet. 

Sensory analysis describes more of the variance in two dimensions, each attribute seems to be relevant and better interpretation can be derived from the sensory PCA 



```{r echo=TRUE, eval=TRUE, tidy=TRUE}
wine_std <- read.table("data/standardized.csv",header=TRUE, sep=",", na.strings="NA", dec=".", strip.white=TRUE)
HClust.1 <- hclust(dist(model.matrix(~-1 + X1.carmel+X2.spice+X3.veg+X4.oak+X5.hot+X6.floral+X7.chem+X8.trop+X9.citrus+X10.sweet+X11.sour+X12.bitter+X13.astrMF+X14.hotMF+X15.viscMF,wine_std)) , method= "ward")
plot(HClust.1,main="Cluster Dendogram based on Sensory", xlab="Different wines")
```

By clustering the set of wines only using Sensory data, three main groups are formed. Group 1 is a Chardonnay group with heavy oak influence (wines 1,2,3 and 16 are all Chardonnays, no wonder!)
Group 2 has a mix of the Sauvignon blanc, Pinot Grigio and Chardonnay that has no oak any other wine that had a “green” taste to them
Group three has the Riesling and Viognier with the tropical and floral taste 

```{r echo=TRUE, eval=TRUE, tidy=TRUE}
HClust.2 <- hclust(dist(model.matrix(~-1 +        X16.viscosity+X17.density+X18.ethanol+X19.phenolics+X20.Citrate+X21.Tartrate+X22.Malate+X23.Succinate+X24.Lactate+X25.Acetic+X26.Total.Sugars+X27.Glycerol+X28.P+X29.K+X30.Ca+X31.Mg+X32.Na+X33.Cl+X34.pH+X35.TA,
                                                  wine_std)) , method= "ward")
plot(HClust.2,main="Cluster Dendogram based on Chemistry", xlab="Different wines")
```

## Including Plots

You can also embed plots, for example:

```{r echo=TRUE, eval=TRUE, tidy=TRUE}
HClust.3 <- hclust(dist(model.matrix(~-1 +        X1.carmel+X2.spice+X3.veg+X4.oak+X5.hot+X6.floral+X7.chem+X8.trop+X9.citrus+X10.sweet+X11.sour+X12.bitter+X13.astrMF+X14.hotMF+X15.viscMF+X16.viscosity+X17.density+X18.ethanol+X19.phenolics+X20.Citrate+X21.Tartrate+X22.Malate+X23.Succinate+X24.Lactate+X25.Acetic+X26.Total.Sugars+X27.Glycerol+X28.P+X29.K+X30.Ca+X31.Mg+X32.Na+X33.Cl+X34.pH+X35.TA,
                                                  wine_std)) , method= "ward")
plot(HClust.3,main="Cluster Dendogram based on all data", xlab="Different wines")
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
